Effects of the body force on the pedestrian and the evacuation dynamics

### 2020-02-12 paper.tex
correcciones de dorso en la parte de intro (las primeras)

The existence of a ``body force'' in the context of highly dense crowds (say, 5 
to 10$\,$people/m$^2$) is a commonsense matter \cite{henein_2007,fruin_1993}. 
Researchers, however, question the numerical setting for this force in 
the SFM context \cite{lakoba_2005}. As a matter of fact, the usual setting by 
Helbing prevents the excessive overlapping among pedestrians, but it is known to 
accomplish artificially high force levels 
\cite{helbing_2000,lakoba_2005,langston_2006,lin_2017}. The force estimates 
from the SFM appear to be remarkably higher with respect to the reported real 
life data (say, an order of magnitude). The crowd motion simulations, however, 
present quite realistic results \cite{lakoba_2005,langston_2006,dorso_2017}. 
The 
point seems to be that the SFM focuses on the ``collective behavior'' due 
to clogging, missing the ``individualistic'' perspective  of single pedestrians 
or very small groups \cite{helbing_2000,henein_2007,narain_2009}.  \\

Many researchers realized that modifying the SFM may (partially) surpass the 
misleading parameter setting. It was proposed that the pedestrians' 
psychological force (say, the ``social force'') should be suppressed in the 
context of highly dense crowds 
\cite{pelechano_2007,moussaid_2011,alonso_2014,bottinelli_2017}, or smoothly 
quenched according to the crowd density \cite{song_2019}. The authors in 
Refs.~\cite{kabalan_2017,jebrane_2019} further proposed a rigid body model in 
order to completely avoid the overlapping phenomenon. This perspective 
dismisses 
any connection to a ``sliding friction''. Conversely, other authors tried to 
limit the pedestrians acceleration by introducing ``static friction'' between 
the pedestrians and the floor \cite{wang_2019}. This kind of friction, however, 
reduces the effective willings of the pedestrians.  \\

A unique (and universal) setting appears not available yet to
our knowledge. 
 The reason is that different numerical settings can lead to 
the same crowd dynamics. Actually, only a small set of dimensionless 
``numbers'' control the crowd dynamics \cite{dorso_2019}. These are similar to 
those encountered in other active matter systems (P\'eclet number, etc.)  
\cite{marchetti_2014}. We may hypothesize that while the dimensionless 
``numbers'' provide some kind of control on the 
collective behavior in 
crowds, only a few numerical settings can attain an ``individualistic'' 
meaning. 
 \\

The numerical setting for the ``faster is slower'' effect presented by 
Helbing and co-workers is somewhat cumbersome 
Ref.~\cite{helbing_2000,dorso_2017,dorso_2019}. Although a single 
parameter (say, the desired velocity $v_d$) is numerically varied to 
explain the phenomenon, the researcher looses sight of the dimensionless 
``numbers'' that truly control the crowd dynamics. The setting in 
Ref.~\cite{helbing_2000} also misses the ``faster is faster'' effect reported 
to occur at very high pedestrian densities \cite{dorso_2017,haghani_2019}. 
Alternatively, the empirical fundamental diagram raises as a point of reference 
for the SFM control parameters \cite{helbing_2007,dorso_2017}. \\

The fundamental diagram exhibits the flux behavior for either low dense crowds 
(with seldom contacts between pedestrians) and highly dense crowds (dominated 
by 
body forces and sliding friction). The latter usually experiences a flux 
slowing 
down, but other behaviors are also possible \cite{helbing_2007,lohner_2018}. In 
light of our previous hypothesis, we may suspect that the  modeling of the 
``flux slowing down'' within the context of the SFM will require the proper 
setting of the (dimensionless) controlling parameters. We examined this working 
hypothesis in Ref.~\cite{dorso_2019}, but limiting the parameter exploration to 
the sliding friction, disregarding the body force. \\

>>>>

The existence of a ``body force'' in the context of highly dense crowds (say, 5 
to 10$\,$people/m$^2$) is a commonsense matter \cite{henein_2007,fruin_1993}. 
Researchers, however, question the numerical setting for this force in 
the SFM context \cite{lakoba_2005}. As a matter of fact, the usual 
\textcolor{blue}{\sout{setting by} set of parameters provided by }
Helbing prevents the excessive overlapping among pedestrians, but it is known to 
accomplish artificially high force levels 
\cite{helbing_2000,lakoba_2005,langston_2006,lin_2017}. The force estimates 
from the SFM appear to be remarkably higher with respect to the reported real 
life data (say, an order of magnitude). The crowd motion simulations, however, 
present \textcolor{blue}{\sout{quite}} realistic results \cite{lakoba_2005,langston_2006,dorso_2017}. 
The 
point seems to be that the SFM focuses on the ``collective behavior'' due 
to clogging, missing the ``individualistic'' perspective  of single pedestrians 
or very small groups \cite{helbing_2000,henein_2007,narain_2009}.  \\

Many researchers realized that modifying the SFM may (partially)
\textcolor{blue}{\sout{ surpass the misleading parameter setting}
fix the misleading results}. It was proposed that the pedestrians' 
psychological force (say, the ``social force'') should be suppressed in the 
context of highly dense crowds 
\cite{pelechano_2007,moussaid_2011,alonso_2014,bottinelli_2017}, or smoothly 
quenched according to the crowd density \cite{song_2019}. The authors in 
Refs.~\cite{kabalan_2017,jebrane_2019} further proposed a rigid body model in 
order to completely avoid the overlapping phenomenon. This perspective 
dismisses 
any connection to a ``sliding friction''. Conversely, other authors tried to 
limit the pedestrians acceleration by introducing ``static friction'' between 
the pedestrians and the floor \cite{wang_2019}. This kind of friction, however, 
reduces the effective willings of the pedestrians.  \\

A unique (and universal) \textcolor{blue}{set of parameters} appears not
available yet to our knowledge. 
 The reason is that different numerical \textcolor{blue}{set of parameters} can lead to 
the same crowd dynamics. Actually, only a small set of dimensionless 
``numbers'' control the crowd dynamics \cite{dorso_2019}. These are similar to 
those encountered in other active matter systems (P\'eclet number, etc.)  
\cite{marchetti_2014}. We may hypothesize that while the dimensionless 
\textcolor{blue}{parameters} provide some kind of control on the 
collective behavior in 
crowds, only a few numerical settings can attain an ``individualistic'' 
meaning. 
 \\

The numerical \textcolor{blue}{parameters} for the ``faster is slower'' effect presented by 
Helbing and co-workers is somewhat cumbersome 
Ref.~\cite{helbing_2000,dorso_2017,dorso_2019}. Although a single 
parameter (say, the desired velocity $v_d$) is numerically varied to 
explain the phenomenon, the researcher looses sight of the dimensionless 
``numbers'' that truly control the crowd dynamics. The \textcolor{blue}{set of parameters} in 
Ref.~\cite{helbing_2000} also misses the ``faster is faster'' effect reported 
to occur at very high pedestrian densities \cite{dorso_2017,haghani_2019}. 
\textcolor{blue}{Moreover}, the empirical fundamental diagram raises as a point of reference 
for the SFM control parameters \cite{helbing_2007,dorso_2017}. \\

The fundamental diagram exhibits the flux behavior for either low \textcolor{blue}{density} crowds 
(with \textcolor{blue}{rare} contacts between pedestrians) and highly dense crowds (dominated 
by body forces and sliding friction). The latter usually experiences a flux 
slowing 
down, but other behaviors are also possible \cite{helbing_2007,lohner_2018}. In 
light of our previous hypothesis, we may suspect that the  modeling of the 
``flux slowing down'' within the context of the SFM will require the proper 
setting of the (dimensionless) controlling parameters. We examined this working 
hypothesis in Ref.~\cite{dorso_2019}, but limiting the parameter exploration to 
the sliding friction, disregarding the body force. \\
